### 2004-07-26 Documentation/Manual.tex
last changes am Manual

\newcommand{\capdef}{}
\newcommand{\mycaption}[2][\capdef]{\renewcommand{\capdef}{#2}%
        \caption[#1]{{ %\itshape
 {\footnotesize #2}}}}
\makeatletter
\renewcommand{\fnum@table}{\textbf{\tablename~\thetable}}
\renewcommand{\fnum@figure}{\textbf{\figurename~\thefigure}}
\makeatother

@@

\begin{center}
{ \Large \bf
{\Huge \GLOBES\ 2.0} \\
General Long Baseline Experiment Simulator \\ }
\vspace*{0.5cm}
{\large \bf User's and experiment definition manual }

@@

Current affiliation(s):

\vspace*{0.5cm}

{\em $^{\mathrm{a},\mathrm{b},\mathrm{c}}$%
       Institut f\"ur Theoretische Physik, Physik--Department,\\
       Technische Universit\"at M\"unchen,
       James--Franck--Strasse, D--85748 Garching, Germany}

@@

\begin{center}
Version from \today
\end{center}

@@

\GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
software package to simulate neutrino oscillation 
long baseline and reactor experiments. On the
one hand, it contains a comprehensive abstract experiment definition
language (AEDL), which allows to describe most classes 
of long baseline experiments
at an abstract level. On the other hand, it provides a C-library to 
process the experiment information in order to obtain oscillation
probabilities, rate vectors, and $\Delta \chi^2$-values. Currently, 
\GLOBES\ is available for GNU/Linux. Since the source code is included,
the port to other operating systems is in principle possible.

@@

\GLOBES\ is developed for academic use. Thus, the \GLOBES\ Team would
appreciate being given academic credit for it. Whenever you use \GLOBES\
to produce a publication or a talk indicate that you have used \GLOBES\ and
please cite the reference~\cite{globes_paper}
\begin{quote}
P. Huber, M. Lindner and W. Winter ... 
\end{quote}
but \emph{not} this manual. This manual itself is not scientific 
publication and will not be submitted to a scientific journal. 
It will evolve during time since it is intended for 
regular revision. Besides that, many of the data which are used by \GLOBES\ and distributed together with it should be properly referenced. 
For details see below.

@@

It is recommended that you use the same style for your own input files, since, in case they are distributed, everbody will know how to correctly reference your work.

@@

\paragraph{Note:} All examples for application software in C do require
a C++ compiler to be properly compiled. For pedagogical reasons, variable
declarations are done at that place where the variable is needed for the
first time, which is at variance with C syntax but not with C++ syntax.
That is the only way in which the examples deviate form ISO C.

%Finally, \Part~\ref{part:3} of this manual serves as reference manual,
%and the appendix contains further supplementary information.

@@

We would like to thank Martin Freund, who wrote the very first
version of a three-flavour matter profile treatment many years ago.
PH is especially thankful for the invaluable advice of Thomas Fischbacher on
many design issues. 
In addition, we would like to thank
Mark Rolinec for his help to translate the experiment descriptions
into \AEDL , and for creating the illustrations in this manuscript.
Finally, thanks to all the people who have been pushing this project
for many years, to the ones who have been continuing asking for the 
publication of the software, and the referees of several of our
papers for suggestions which lead to improvements in the software.

@@

\addcontentsline{toc}{chapter}{Bibliography}
\bibliographystyle{./unsrt}
\bibliography{references}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\newcommand{\glbxxx}{hep-ph/0407xxx}
\newcommand{\capdef}{}
\newcommand{\mycaption}[2][\capdef]{\renewcommand{\capdef}{#2}%
        \caption[#1]{{ %\itshape
 {\footnotesize #2}}}}
\makeatletter
\renewcommand{\fnum@table}{\textbf{\tablename~\thetable}}
\renewcommand{\fnum@figure}{\textbf{\figurename~\thefigure}}
\makeatother

@@

\begin{center}
{ \Large \bf
{\Huge \GLOBES} \\
General Long Baseline Experiment Simulator \\ }
\vspace*{0.5cm}
{\large \bf User's and experiment definition manual }

@@

@@

\begin{center}
Version from \today\ for \GLOBES\ 2.0
\end{center}

\vspace*{0.5cm}
\vfill
\hrule

{\em\small $^{\mathrm{a},\mathrm{b},\mathrm{c}}$%
       Institut f\"ur Theoretische Physik, Physik--Department,\\
       Technische Universit\"at M\"unchen,
       James--Franck--Strasse, D--85748 Garching, Germany}

@@

\vspace*{-2ex}
\GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
software package to simulate neutrino oscillation 
long baseline and reactor experiments. On the
one hand, it contains a comprehensive abstract experiment definition
language (AEDL), which allows to describe most classes 
of long baseline experiments
at an abstract level. On the other hand, it provides a C-library to 
process the experiment information in order to obtain oscillation
probabilities, rate vectors, and $\Delta \chi^2$-values. Currently, 
\GLOBES\ is available for GNU/Linux. Since the source code is included,
the port to other operating systems is in principle possible.

@@

\GLOBES\ is developed for academic use. Thus, the \GLOBES\ Team would
appreciate being given academic credit for it. Whenever you use \GLOBES\
to produce a publication or a talk indicate that you have used \GLOBES\ and
please cite the reference~\cite{globes_paper}
\begin{quote}
P. Huber, M. Lindner and W. Winter\\
Simulation of long baseline neutrino oscillation experiments with \GLOBES\\
arXiv:\glbxxx.  
\end{quote}
but \emph{not} this manual. This manual itself is not a scientific 
publication and will not be submitted to a scientific journal. 
It will evolve during time since it is intended for 
regular revision. Besides that, many of the data which are used by \GLOBES\ 
and distributed together with it should be properly referenced. 
For details see below.

@@

It is recommended that you use the same style for your own input files, since, in case they are distributed, everybody will know how to correctly reference your work.

@@

\paragraph{Note:} All examples for application software in C do require
a C++ compiler to be properly compiled. For pedagogical reasons, variable
declarations are done at that place where the variable is needed for the
first time, which is at variance with C syntax but not with C++ syntax.
That is the only way in which the examples deviate form ISO C. Moreover
the actual numerical values of the results of the examples may be different
from the ones in this manual.

@@

We would like to thank Martin Freund, who wrote the very first
version of a three-flavour matter profile treatment many years ago.
PH is especially thankful for the invaluable advice of Thomas Fischbacher on
many design issues. 
In addition, we would like to thank
Mark Rolinec for his help to translate the experiment descriptions
into \AEDL, and for creating the illustrations in this manuscript.
Finally, thanks to all the people who have been pushing this project
for many years, to the ones who have been continuing asking for the 
publication of the software, and the referees of several of our
papers for suggestions which lead to improvements in the software.

This work and the development of \GLOBES\ have been supported by:
\begin{itemize}
\item Physik Department der Technischen Universit\"at M\"unchen
\item Max-Planck-Institut f\"ur Physik 
\item Sonderforschungsbereich 375 f\"ur Astro-Teilchenphysik\\ 
der Deutschen Forschungsgemeinschaft
\item Studienstiftung des Deutschen Volkes 
\end{itemize}

@@

\addcontentsline{toc}{chapter}{Bibliography}
\bibliographystyle{./OurBibTeX}
\bibliography{references}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2004-07-26 Documentation/part1.tex
last changes am Manual

Let us now locate the exact position of the sgn-degeneracy:

@@

We will in this part not go into details of the experiment
definition. The pre-defined experiment prototypes in the {\tt data}
subdirectory are summarized in \tabl{experiments}. They correspond
(except from minor modifications) to the experiments in the
respective references in the table.

@@

\underline{$\beta$-Beams:} & \multicolumn{3}{l}{In preparation} \\[0.1cm]
\hline
\end{tabular}
\end{center}
\mycaption{\label{tab:experiments} \index{norm}{Experiment files (table)}
\index{norm}{glb files@{\tt glb}-files} 
Pre-defined experiment prototypes, their filenames (to be used in \GLB{InitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
different implementation of the energy threshold function). Note that
some of these experiments are outdated in terms of integrated luminosities,
baseline, fluxes, efficiencies, or other factors.}
\end{table}

@@

@@

\example{Finding the $\mathrm{sgn}(\ldm)$-degeneracy}{
\label{ex:sgndeg}
\index{norm}{Degeneracies!sgn($\ldm$)-degeneracy}

@@

The example on page~\pageref{ex:sgndeg} illustrates how to locate the $\mathrm{sgn}(\ldm)$-degeneracy and show the corresponding degenerate solution in the $\stheta$-$\deltacp$-plane together with the original solution.
In this case, the position of the degeneracy can be easily guessed to be 
at the true parameter values but with inverted $\ldm$. The minimizer then runs off the plane of these parameters into the local minimum. It is very important to take into account the position of the degeneracy off this plane, since the actual $\chi^2$ in the minimum is probably lower than in the plane. Thus, the degeneracy may not even appear at the chosen confidence level in the plane, but it does appear at the real minimum. The two sections through the fit manifold shown in the figure on page~\pageref{ex:sgndeg} therefore do not appear at the same oscillation parameter values (except from the ones shown in the 
figure).\footnote{The discussed figure on page~\pageref{ex:sgndeg} is produced by  {\tt glbChiSys} and thus only represents a section through the fit manifold. For the projection including correlations, one may rather want to use {\tt glbChiThetaDelta}.}

>>>>

Let us now locate the exact position\footnote{For a exact definition of 
inverted hierarchy, see page~\pageref{mass_ordering}.} of the sgn-degeneracy:

@@

We will in this part not go into details of the experiment
definition. The pre-defined experiment prototypes in the {\tt data}
subdirectory are summarized in \tabl{experiments}. They correspond
(except from minor modifications) to the experiments in the
respective references in the table. These files are installed to
the directory \verb+${prefix}/share/globes+ which usually defaults
to  \verb+/usr/local/share/globes+. It is useful to add this path
to the value of \verb+GLB_PATH+.
\index{norm}{glb files@{\tt glb}-files!installation}

@@

\underline{$\beta$-Beams:} & \multicolumn{3}{l}{In preparation} \\[0.1cm]
\hline
\end{tabular}
\end{center}
\mycaption{\label{tab:experiments} \index{norm}{Experiment files (table)}
\index{norm}{glb files@{\tt glb}-files} 
Pre-defined experiment prototypes, their filenames (to be used in \GLB{InitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
different implementation of the energy threshold function). Note that
some of these experiments are outdated in terms of integrated luminosities,
baseline, fluxes, efficiencies, or other factors. In any case these file are installed along with \GLOBES.}
\end{table}

@@

\label{mass_ordering}
\index{norm}{Mass hierarchy}
\paragraph{Note:} Inverting the mass hierarchy is not
precisely the same than to change from $\Delta m^2_{31}
\rightarrow -\Delta m^2_{31}$.
In this case the absolute value of $\Delta m^2_{32}$ changes also, which
introduces a new frequency to the problem. Therefore, if we assume
normal hierarchy  whenever $\Delta m^2_{31}>0$, the corresponding
point in parameters space for inverted hierarchy is given by $\Delta 
m^2_{31}\rightarrow - \Delta m^2_{31} + \Delta m^2_{21}$, because
with this defintion the absolute value of  $\Delta m^2_{32}$ is unchanged
and no new frequency is introduced.\\

@@

\example{Finding the $\mathrm{sgn}(\ldm)$-degeneracy}{
\label{ex:sgndeg}
\index{norm}{Degeneracies!sgn($\ldm$)-degeneracy}
\index{norm}{Mass hierarchy}

@@

The example on page~\pageref{ex:sgndeg} illustrates how to locate the $\mathrm{sgn}(\ldm)$-degeneracy and show the corresponding degenerate solution in the $\stheta$-$\deltacp$-plane together with the original solution.
In this case, the position of the degeneracy can be easily guessed to be 
at the true parameter values but with inverted\footnote{For a exact 
definition of 
inverted hierarchy, see page~\pageref{mass_ordering}.} $\ldm$. The minimizer then runs off the plane of these parameters into the local minimum. It is very important to take into account the position of the degeneracy off this plane, since the actual $\chi^2$ in the minimum is probably lower than in the plane. Thus, the degeneracy may not even appear at the chosen confidence level in the plane, but it does appear at the real minimum. The two sections through the fit manifold shown in the figure on page~\pageref{ex:sgndeg} therefore do not appear at the same oscillation parameter values (except from the ones shown in the 
figure).\footnote{The discussed figure on page~\pageref{ex:sgndeg} is produced by  {\tt glbChiSys} and thus only represents a section through the fit manifold. For the projection including correlations, one may rather want to use {\tt glbChiThetaDelta}.}
